---
title: SEM II Final Presentation
subtitle: Latent Interaction
author: B Spinner
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true 
    code_folding: hide
    theme: cerulean
--- 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}

library(haven)
library(flexdashboard)
pacman::p_load(haven, flextable, data.table, ggthemes, sandwich, flexdashboard, ggplot2, ggExtra, grid, tidyverse, rio, here, modelsummary, janitor, ggThemeAssist, gghighlight, ggridges)

Latinx <- read_sav(here("data/Latinx Health Behaviors and Health Outcomes dataset_7.14.2020.sav"))

#View(Latinx)

Latinx_sub <- Latinx %>% 
  select("Age_combine", "Hisp_combine", "Gender_combine", "Household_Income_combine", "RelationshipStatus_combine", "PHQ_1_combine","PHQ_2_combine","PHQ_3_combine", "PHQ_4_combine","PHQ_5_combine","PHQ_6_combine", "PHQ_7_combine","PHQ_8_combine","PHQ_9_combine", "SDQ_1_combine", "SDQ_2_combine", "SDQ_3_combine", "SDQ_4_combine", "SDQ_5_combine", "SDQ_6_combine", "SDQ_7_combine", "SDQ_8_combine", "AMAS_AMERICANculture_1_combine", "AMAS_AMERICANculture_2_combine", "AMAS_AMERICANculture_3_combine", "AMAS_AMERICANculture_4_combine", "AMAS_AMERICANculture_5_combine", "AMAS_AMERICANculture_6_combine", "AMAS_NATIVEculture_7_combine", "AMAS_NATIVEculture_8_combine", "AMAS_NATIVEculture_9_combine", "AMAS_NATIVEculture_10_combine", "AMAS_NATIVEculture_11_combine", "AMAS_NATIVEculture_12_combine")

# Dropped PHQs that weren't reltaed to depression specifically: 3, 
# Dropped SDQs that weren't in Likert: 1, 2, 6 (maybe 5) 
Latinx_miss <- Latinx_sub %>% 
  select("Age_combine", "Household_Income_combine", "PHQ_1_combine","PHQ_2_combine", "PHQ_4_combine","PHQ_5_combine","PHQ_6_combine", "PHQ_7_combine","PHQ_8_combine","PHQ_9_combine", "SDQ_3_combine", "SDQ_4_combine", "SDQ_5_combine", "SDQ_7_combine", "SDQ_8_combine", "AMAS_AMERICANculture_1_combine", "AMAS_AMERICANculture_2_combine", "AMAS_AMERICANculture_3_combine", "AMAS_AMERICANculture_4_combine", "AMAS_AMERICANculture_5_combine", "AMAS_AMERICANculture_6_combine", "AMAS_NATIVEculture_7_combine", "AMAS_NATIVEculture_8_combine", "AMAS_NATIVEculture_9_combine", "AMAS_NATIVEculture_10_combine", "AMAS_NATIVEculture_11_combine", "AMAS_NATIVEculture_12_combine")


```

# SEM II Final Project Presentation 
## Utilizing the Product Indicators Method in testing for the moderating effect of acculturation in the relationship between latent sleep and depression adjusting for age and household income (from SEM I) among Latinx adults in the US. 

## Introduction
Poor sleep significantly impacts mental and physical health, particularly contributing to conditions like depression, anxiety, and severe psychological distress (SPD). We previously explored how poor sleep/sleep disorders elevated the risk of SPD among a diverse Hispanic/Latino(a) adult population (N = 798). The present study examines how acculturation moderates the relationship between sleep disorders and depression among US Latinx adults. Acculturation stress and ethnic discrimination are associated with poor sleep in Hispanics/Latinos (Alcántara et al., 2017). However, the prevalence and understanding of sleep disorders among US Latinx adults is not well known (Loredo et al., 2010). Using a product indicators method in SEM, we found that the introduction of acculturation as a moderating variable significantly altered the direct effect of sleep on depression, which became non-significant (estimate = 0.031, p = 0.514). The interaction term between sleep disorders and acculturation was significant (estimate = 0.093, p < .001), indicating that the effect of sleep on depression varies across different levels of acculturation (sense of belonging in US).

## Research Question 
When controlling for covariates, does acculturation moderate the relationship between poor sleep and depression, on average among US Latinx adults?

## Hypothesis
When controlling for _Age_ and _Household Income_ we predict that Latinx adults in the US with sleep disorders will, on average, have increased symptoms of depression when they are experiencing higher levels of acculturation.  Specifically, when levels of acculturation are high, they will have  higher scores on the sleep disorders questionnaire (SDQ), as level of depression increases (score on the PhQ). 

## Mardia's test for Bivariate relationship 
## Missingness evaluated usign MCAR test
There is no significant evidence against the MCAR assumption based on the test performed (p> .05). And data do not meet multivariate normality assumptions according to Mardia's test results (p=0). 

Moving forward in analyses we will be adjusting for levels of missingness using Full Information Maximum Likelihood (FIML) and the Maximum Likelihood Estimation with Robust Standard Errors (“MLR”) to adjust for levels of non-multivariate normality.

```{r}
library(naniar)
mcar_test(Latinx_miss)

library(MVN)       


mvn(Latinx_miss, mvnTest = "mardia")
```

### Relationship between latent sleep and depression adjusting for age and household income (from SEM I)

We previously found a significant positive association between poorer sleep and depression among Latinx adults in the US, on average, adjusting for age and household income. 

```{r}
library(lavaan)
library(lavaanPlot)
library(psych)

# dont use SDQ 5 it's negative, SDQ 3 and 4 also such .2 or .3 factor loadings. these questions ask about what you were doing while sleeping. not many of us know. so just a poor measure. 

LVsem <-  ' latent_sleep =~ SDQ_3_combine + SDQ_4_combine + SDQ_7_combine + SDQ_8_combine 

latent_depress =~ PHQ_1_combine + PHQ_2_combine + PHQ_4_combine + PHQ_5_combine +  PHQ_6_combine + PHQ_7_combine + PHQ_8_combine + PHQ_9_combine

latent_depress ~ latent_sleep + Age_combine + Household_Income_combine 

'
out1 <- sem(LVsem, missing = "fiml", estimator = "MLR", data=Latinx_miss)

summary(out1, stand=TRUE, fit.measures=TRUE)

#modelsummary::modelsummary(out1)
# library(sandwich)
# library(clubSandwich)
omit_rows <- c(1,2,3,4,5,6,7,8,9,10,11, 12,14, 15, 16)
include_rows <- c("SEM Model Fit")
modelsummary::modelsummary(list( "SEM Model Fit" = out1),
                            stars=T,
              fmt=3,
             gof_omit = c("Adj.|Log.Lik."),
             coef_omit = omit_rows, 
              col_names = include_rows,
             coef_rename = c("latent_depress ~ latent_sleep" = "LV Depression and LV Sleep",
                             "latent_depress ~ Age_combine" = "LV Depression and Age",
                             "latent_depress ~ Household_Income_combine" = "LV Depression and Household Income"),
                             notes = c("Notes: Table presents coefficients estimated with maximum likelihood estimation with robust standard errors (MLR). Missing values were addressed using Full Information Maximum Likelihood (FIML) estimation."))
            # output = "figures/model_summary.png")
#install.packages("webshot2")

lavaanPlot(model=out1, 
           stand=TRUE, 
           digits = 3, 
           coefs=TRUE, 
           covs=TRUE, 
           graph_options = list(rankdir = "TB"),  
           stars=c("regress"))


# ggsave("figures/figure_sem.png", plot = fig)
# export_graph(fig, file = "figures/figure_sem.svg")



library(semPlot)
 
fig1<-semPaths(out1,                   
         whatLabels = "stand", 
         layout = "tree2",
         rotation = 4,
         nCharNodes = 0,             
         edge.width = 2.5,          
         edge.label.cex = 1,
         exoCov = FALSE,
         intercepts=FALSE,           
         residuals=FALSE,         
         equalizeManifests = TRUE,  
         sizeLat=10,                
         sizeMan=8,
         unCol = "#070b8c")


```

## Fit indices for Bivariate relationship 

```{r}
### Modified Fit and Rsquare
fit_indices<-fitMeasures(out1,c("chisq", "df", "pvalue", "cfi", "rmsea"))


fit_indices<-as.data.frame(fit_indices)
#View(fit_indices)
reportfits<- as.data.frame(t(fit_indices)) # transpose all but the first column (name)

#View(reportfits)

reportfits$chisq_df<-(reportfits[,c('chisq')]/reportfits[,c('df')])
    
 
reportfits<-as.data.frame(reportfits)

R_Square<-lavInspect(out1,"r-square")
R_Square<-as.data.frame(R_Square)

reportRsq<- as.data.frame(t(R_Square)) # transpose all but the first column (name)
#View(reportRsq)
# 
# LV_Rsq<-reportRsq %>% 
#   select(c("out3"))


library(kableExtra)
getOption("digits")

reportfits%>%
  kbl(col.names = c ("chisq", "df", "pvalue", "cfi", "rmsea", "chisq_df"), digits = 3) %>% 
  kable_classic(full_width = F, html_font = "Cambria")

```


```{r, include=FALSE}
LVsem_accult <-  ' latent_accult =~ AMAS_AMERICANculture_1_combine + AMAS_AMERICANculture_2_combine + AMAS_AMERICANculture_3_combine + AMAS_AMERICANculture_4_combine + AMAS_AMERICANculture_5_combine + AMAS_AMERICANculture_6_combine 

latent_depress =~ PHQ_1_combine + PHQ_2_combine + PHQ_4_combine + PHQ_5_combine +  PHQ_6_combine + PHQ_7_combine + PHQ_8_combine + PHQ_9_combine

latent_depress ~ latent_accult + Age_combine + Household_Income_combine 

'
out2 <- sem(LVsem_accult, missing = "fiml", estimator = "MLR", data=Latinx_miss)

summary(out2, stand=TRUE, fit.measures=TRUE)

modelsummary::modelsummary(out2)
# library(sandwich)
# library(clubSandwich)
# omit_rows <- c(1,2,3,4,5,6,7,8,9,10,14)
# include_rows <- c("SEM Model Fit")
# modelsummary::modelsummary(list( "SEM Model Fit" = out2),
#                             stars=T,
#               fmt=3,
#              gof_omit = c("Adj.|Log.Lik."),
#              coef_omit = omit_rows, 
#               col_names = include_rows,
#              coef_rename = c("latent_depress ~ latent_sleep" = "LV Depression and LV Sleep",
#                              "latent_depress ~ Age_combine" = "LV Depression and Age",
#                              "latent_depress ~ Household_Income_combine" = "LV Depression and Household Income"),
#                              notes = c("Notes: Table presents coefficients estimated with maximum likelihood estimation with robust standard errors (MLR). Missing values were addressed using Full Information Maximum Likelihood (FIML) estimation."))
#             # output = "figures/model_summary.png")
# #install.packages("webshot2")

lavaanPlot(model=out2, 
           stand=TRUE, 
           digits = 3, 
           coefs=TRUE, 
           covs=TRUE, 
           graph_options = list(rankdir = "TB"),  
           stars=c("regress"))


# ggsave("figures/figure_sem.png", plot = fig)
# export_graph(fig, file = "figures/figure_sem.svg")



library(semPlot)
 
fig2<-semPaths(out2,                   
         whatLabels = "stand", 
         layout = "tree2",
         rotation = 4,
         nCharNodes = 0,             
         edge.width = 2.5,          
         edge.label.cex = 1,
         exoCov = FALSE,
         intercepts=FALSE,           
         residuals=FALSE,         
         equalizeManifests = TRUE,  
         sizeLat=10,                
         sizeMan=8,
         unCol = "#070b8c")

# Latent accult does not significantly predict depression. it does sig predict sleep. 
```

```{r, include=FALSE}
LVsem_accult_sleep <-  ' latent_accult =~ AMAS_AMERICANculture_1_combine + AMAS_AMERICANculture_2_combine + AMAS_AMERICANculture_3_combine + AMAS_AMERICANculture_4_combine + AMAS_AMERICANculture_5_combine + AMAS_AMERICANculture_6_combine 
latent_sleep =~ SDQ_3_combine + SDQ_4_combine + SDQ_7_combine + SDQ_8_combine 

latent_sleep ~ latent_accult + Age_combine + Household_Income_combine '

out3 <- sem(LVsem_accult_sleep, missing = "fiml", estimator = "MLR", data=Latinx_miss)

summary(out3, stand=TRUE, fit.measures=TRUE)

modelsummary::modelsummary(out3)
# library(sandwich)
# library(clubSandwich)
# omit_rows <- c(1,2,3,4,5,6,7,8,9,10,14)
# include_rows <- c("SEM Model Fit")
# modelsummary::modelsummary(list( "SEM Model Fit" = out2),
#                             stars=T,
#               fmt=3,
#              gof_omit = c("Adj.|Log.Lik."),
#              coef_omit = omit_rows, 
#               col_names = include_rows,
#              coef_rename = c("latent_depress ~ latent_sleep" = "LV Depression and LV Sleep",
#                              "latent_depress ~ Age_combine" = "LV Depression and Age",
#                              "latent_depress ~ Household_Income_combine" = "LV Depression and Household Income"),
#                              notes = c("Notes: Table presents coefficients estimated with maximum likelihood estimation with robust standard errors (MLR). Missing values were addressed using Full Information Maximum Likelihood (FIML) estimation."))
#             # output = "figures/model_summary.png")
# #install.packages("webshot2")

lavaanPlot(model=out3, 
           stand=TRUE, 
           digits = 3, 
           coefs=TRUE, 
           covs=TRUE, 
           graph_options = list(rankdir = "TB"),  
           stars=c("regress"))


# ggsave("figures/figure_sem.png", plot = fig)
# export_graph(fig, file = "figures/figure_sem.svg")



library(semPlot)
 
fig3<-semPaths(out3,                   
         whatLabels = "stand", 
         layout = "tree2",
         rotation = 4,
         nCharNodes = 0,             
         edge.width = 2.5,          
         edge.label.cex = 1,
         exoCov = FALSE,
         intercepts=FALSE,           
         residuals=FALSE,         
         equalizeManifests = TRUE,  
         sizeLat=10,                
         sizeMan=8,
         unCol = "#070b8c")


```


# Product Indicators Method
## Testing for interaction in bivariate relationship using latent acculturation variable

	•	Acculturation (latent_accult) has a significant negative effect on depression.
	•	The interaction term (latent_int) shows a significant moderation effect, indicating that the relationship between sleep and depression is influenced by the level of acculturation.
	•	Age and Household Income both have significant negative effects on depression.
	•	After the introduction of our moderating variable, Sleep (latent_sleep) no longer has a significant direct effect on depression.

**Conclusion**
The interaction term (latent_int) being significant indicates that acculturation significantly alters how sleep impacts depression. For instance, sleep might have a stronger or weaker effect on depression at different levels of acculturation. In conclusion, the impact of sleep on depression varies with different levels of acculturation, emphasizing the importance of considering acculturation in the context of mental health research.

```{r}
library(lavaan)
library(lavaanPlot)
library(semTools)

Latinx_miss$Interact1 = Latinx_miss$SDQ_7_combine * Latinx_miss$AMAS_AMERICANculture_1_combine
Latinx_miss$Interact2 = Latinx_miss$SDQ_8_combine * Latinx_miss$AMAS_AMERICANculture_1_combine
Latinx_miss$Interact3 = Latinx_miss$SDQ_7_combine * Latinx_miss$AMAS_AMERICANculture_2_combine
Latinx_miss$Interact4 = Latinx_miss$SDQ_8_combine * Latinx_miss$AMAS_AMERICANculture_2_combine
Latinx_miss$Interact5 = Latinx_miss$SDQ_7_combine * Latinx_miss$AMAS_AMERICANculture_3_combine
Latinx_miss$Interact6 = Latinx_miss$SDQ_8_combine * Latinx_miss$AMAS_AMERICANculture_3_combine
Latinx_miss$Interact7 = Latinx_miss$SDQ_7_combine * Latinx_miss$AMAS_AMERICANculture_4_combine
Latinx_miss$Interact8 = Latinx_miss$SDQ_8_combine * Latinx_miss$AMAS_AMERICANculture_4_combine
Latinx_miss$Interact9 = Latinx_miss$SDQ_7_combine * Latinx_miss$AMAS_AMERICANculture_5_combine
Latinx_miss$Interact10 = Latinx_miss$SDQ_8_combine * Latinx_miss$AMAS_AMERICANculture_5_combine
Latinx_miss$Interact11 = Latinx_miss$SDQ_7_combine * Latinx_miss$AMAS_AMERICANculture_6_combine
Latinx_miss$Interact12 = Latinx_miss$SDQ_8_combine * Latinx_miss$AMAS_AMERICANculture_6_combine

Intmod <- "

latent_depress =~ PHQ_1_combine + PHQ_2_combine + PHQ_4_combine + PHQ_5_combine +  PHQ_6_combine + PHQ_7_combine + PHQ_8_combine + PHQ_9_combine

latent_sleep =~ SDQ_7_combine + SDQ_8_combine 

latent_accult =~ AMAS_AMERICANculture_1_combine + AMAS_AMERICANculture_2_combine + AMAS_AMERICANculture_3_combine + AMAS_AMERICANculture_4_combine + AMAS_AMERICANculture_5_combine + AMAS_AMERICANculture_6_combine 

latent_int =~ Interact1 + Interact2 + Interact3 + Interact4 + Interact5 + Interact6 + Interact7 + Interact8 + Interact9 + Interact10 + Interact11 + Interact12

latent_depress ~ latent_sleep + latent_accult + latent_int + Age_combine + Household_Income_combine
 "

fitmoder<-sem(model=Intmod, data=Latinx_miss, missing = "FIML", estimator = "MLR")                       

summary(fitmoder,standardized=TRUE)                   
 
lavaanPlot(model=fitmoder,stand=TRUE, 
           node_options = list(shape="box",fontname="Helvetica",fontsize="12"),
           edge_options = list(color="black"),
           graph_options = list(rankdir = "LR"),
           coefs=TRUE, 
           covs=FALSE,
           stars=c("regress")) 

probe<-probe2WayMC(fitmoder, c("latent_sleep", "latent_accult", "latent_int"), 
                   "latent_depress","latent_accult", 
                   c(-1,0, 1))
probe

plotProbe(probe,c(-2,2),xlab="Latent sleep and accult", ylab="Latent depress")  

#citation("semTools")

```

```{r}
library(modelsummary)
modelsummary::modelsummary(fitmoder)
 library(sandwich)
 library(clubSandwich)
omit_rows <- c(1,2,3,4,5,6,7,8,9,10,11, 12,14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28)
include_rows <- c("SEM Interaction Fit")
modelsummary::modelsummary(list( "SEM Interaction Fit" = fitmoder),
                            stars=T,
              fmt=3,
             gof_omit = c("Adj.|Log.Lik."),
             coef_omit = omit_rows,
              col_names = include_rows,
                             notes = c("Notes: Table presents coefficients estimated with maximum likelihood estimation with robust standard errors (MLR). Missing values were addressed using Full Information Maximum Likelihood (FIML) estimation."))


        # coef_rename = c("latent_depress ~ latent_sleep" = "LV Depression and LV Sleep",
        #                      "latent_depress ~ Age_combine" = "LV Depression and Age",
        #                      "latent_depress ~ Household_Income_combine" = "LV Depression and Household Income"),
```


## Fit indices for interaction model

```{r}
### Modified Fit and Rsquare
fit_indices2<-fitMeasures(fitmoder,c("chisq", "df", "pvalue", "cfi", "rmsea"))


fit_indices2<-as.data.frame(fit_indices2)
#View(fit_indices)
reportfits2<- as.data.frame(t(fit_indices2)) # transpose all but the first column (name)

#View(reportfits)

reportfits2$chisq_df2<-(reportfits2[,c('chisq')]/reportfits2[,c('df')])
    
 
reportfits2<-as.data.frame(reportfits2)

R_Square2<-lavInspect(fitmoder,"r-square")
R_Square2<-as.data.frame(R_Square2)

reportRsq2<- as.data.frame(t(R_Square2)) # transpose all but the first column (name)
#View(reportRsq)
# 
# LV_Rsq<-reportRsq %>% 
#   select(c("out3"))


library(kableExtra)
getOption("digits")

reportfits%>%
  kbl(col.names = c ("chisq", "df", "pvalue", "cfi", "rmsea", "chisq_df"), digits = 3) %>% 
  kable_classic(full_width = F, html_font = "Cambria")

```



```{r, include=FALSE}
## Factor score method
library(haven)
# 
# latinx_df <- Latinx_miss %>%
#   mutate(across(where(is.labelled), ~ as_factor(.)))

# Assuming your_data is your dataset
latinx_df <- Latinx_miss %>%
  mutate(across(c(SDQ_7_combine, SDQ_8_combine, PHQ_1_combine, PHQ_2_combine, PHQ_4_combine, 
                  PHQ_5_combine, PHQ_6_combine, PHQ_7_combine, PHQ_8_combine, PHQ_9_combine, 
                  AMAS_AMERICANculture_1_combine, AMAS_AMERICANculture_2_combine, 
                  AMAS_AMERICANculture_3_combine, AMAS_AMERICANculture_4_combine, 
                  AMAS_AMERICANculture_5_combine, AMAS_AMERICANculture_6_combine), 
                ~ as.numeric(.)))
facscoremod <- "
                     sleep_FAC =~ SDQ_7_combine + SDQ_8_combine 
                  depress_FAC =~ PHQ_1_combine + PHQ_2_combine + PHQ_4_combine + PHQ_5_combine +  PHQ_6_combine + PHQ_7_combine + PHQ_8_combine + PHQ_9_combine 
                     accult_FAC =~ AMAS_AMERICANculture_1_combine + AMAS_AMERICANculture_2_combine + AMAS_AMERICANculture_3_combine + AMAS_AMERICANculture_4_combine + AMAS_AMERICANculture_5_combine + AMAS_AMERICANculture_6_combine  
                  "

facmodfit<-cfa(model=facscoremod, data= latinx_df, missing = "FIML", estimator = "MLR")                       

facscores <- predict(facmodfit)

facscores<-as.data.frame(facscores) 

head(facscores)

# add factor scores to our main data set

latinx_df$sleep_FAC<-facscores$sleep_FAC
latinx_df$depress_FAC<-facscores$depress_FAC
latinx_df$accult_FAC<-facscores$accult_FAC
```

```{r, include=FALSE}
# note we also need to define a product term of the factor scores involved in the interaction

latinx_df$sleepxaccult_FAC <- facscores$sleep_FAC*facscores$accult_FAC
```


```{r, include=FALSE}
# now define the simple factor score model

buffmoder4 <- "depress_FAC ~ sleep_FAC + accult_FAC +  sleepxaccult_FAC"

fitmoder4<-sem(model=buffmoder4, data=latinx_df, missing = "FIML")                       

summary(fitmoder4,standardized=TRUE)                   
 
lavaanPlot(model=fitmoder4,stand=TRUE, 
           node_options = list(shape="box",fontname="Helvetica",fontsize="12"),
           edge_options = list(color="black"),
           graph_options = list(rankdir = "LR"),
           coefs=TRUE, 
           covs=FALSE,
           stars=c("regress")) 


probe<-probe2WayMC(fitmoder4, c("sleep_FAC", "accult_FAC", "sleepxaccult_FAC"), 
                   "depress_FAC", "accult_FAC",
                   c(-1,0, 1))
probe

plotProbe(probe,c(-2,2),xlab="facscore: sleep and accult", ylab="facscore: depression") 
    
```


```{r}
# library(lavaan)
# citation("lavaan")
# 
# ?citation()
```

